---
title: "Practical Machine Learning Project"
output: html_document
---

The goal of this project is to predict the manner in which the people using movement devices did the exercise. This is the "classe" variable in the training set.

##Data Manipulation
This process removes the columns that have a majority of NA and blank values for both the test and training datasets. It also removes the first six columns so we are left with the raw movement data and the "classe" variable that is in the training set.
```{r}
#pml-testing.csv & pml-training.csv files must be in working directory
test_read <- read.csv("pml-testing.csv", na.strings = c("NA","#DIV/0!",""))
testing <- test_read[,colSums(is.na(test_read))==0] #remove columns with NAs
testing[,7:59] <- sapply(testing[,7:59], as.numeric)
test_20 <- testing[,7:59]

train_read <- read.csv("pml-training.csv", na.strings = c("NA","#DIV/0!",""))
training <- train_read[,colSums(is.na(train_read))==0] #remove columns with NAs
training[,7:59] <- sapply(training[,7:59], as.numeric)
training <- training[,7:60]
```

##Prediction Analysis
```{r, warning=FALSE, message=FALSE}
library(caret)
library(rpart)
library(randomForest)
library(e1071)

set.seed(123)
inTrain <- createDataPartition(y=training$classe, p=.75, list=FALSE)
train <- training[inTrain,]
test <- training[-inTrain,]
```
Linear Discriminant Analysis
```{r}
modlda <- train(classe~., data=training, method="lda")
predictLDA <- predict(modlda, test)
```
Recursive Partition
```{r}
modrp <- rpart(classe~., data=training, method="class")
predictRP <- predict(modrp, test, type="class")
```
Random Forest
```{r}
modrf <- randomForest(classe~., data=training)
predictRF <- predict(modrf, test)
```

Test to see which predictive model is most accurate.
```{r}
confusionMatrix(predictLDA, test$classe)
confusionMatrix(predictRP, test$classe)
confusionMatrix(predictRF, test$classe)
```

Since the Random Forest model has the highest accuracy, we'll us it to predict the classes for the initial 20 test cases.
```{r}
predict(modrf, test_20)
```

