---
title: "Practical Machine Learning Project"
output:
  md_document:
    variant: markdown_github

---
#Practical Machine Learning Project
The goal of this project is to predict the manner in which the people using movement devices did the exercise. This is the "classe" variable in the training set.

##Data Manipulation
This process removes the columns that have a majority of NA and blank values for both the test and training datasets. It also removes the first six columns so we are left with the raw movement data and the "classe" variable that is in the training set.
```{r}
#pml-testing.csv & pml-training.csv files must be in working directory
test_read <- read.csv("pml-testing.csv", na.strings = c("NA","#DIV/0!",""))
testing <- test_read[,colSums(is.na(test_read))==0] #remove columns with NAs
testing[,7:59] <- sapply(testing[,7:59], as.numeric)
test_20 <- testing[,7:59]

train_read <- read.csv("pml-training.csv", na.strings = c("NA","#DIV/0!",""))
training <- train_read[,colSums(is.na(train_read))==0] #remove columns with NAs
training[,7:59] <- sapply(training[,7:59], as.numeric)
training <- training[,7:60]
```

##Prediction Analysis
```{r, warning=FALSE, message=FALSE}
library(caret)
library(rpart)
library(randomForest)
library(e1071)

set.seed(123)
inTrain <- createDataPartition(y=training$classe, p=.75, list=FALSE)
train <- training[inTrain,]
test <- training[-inTrain,]
```
Linear Discriminant Analysis
```{r}
modlda <- train(classe~., data=training, method="lda")
predictLDA <- predict(modlda, test)
```
Recursive Partition
```{r}
modrp <- rpart(classe~., data=training, method="class")
predictRP <- predict(modrp, test, type="class")
```
Random Forest
```{r}
modrf <- randomForest(classe~., data=training)
predictRF <- predict(modrf, test)
```

This step uses the confusionMatrix function for cross validation and to calculate the out of sample error for each of the three prediction models.
```{r}
confusionMatrix(predictLDA, test$classe)
confusionMatrix(predictRP, test$classe)
confusionMatrix(predictRF, test$classe)
```

Since the Random Forest model has the highest accuracy of 1, we'll use it to predict the classes for the initial 20 test cases.
```{r}
predict(modrf, test_20)
```

